### 2023-07-12 calculus.tex
Copy over updated symbols

Conditionals, however, present a more interesting case. In the unmarked language, we have both explicit
synthetic and analytic rules:
%
\begin{mathpar}
  \judgment{
    \ctxAnaTypeU{\ctx}{\EMV_1}{\TBool} \\\\
    \ctxSynTypeU{\ctx}{\EMV_2}{\TMV_1} \\
    \ctxSynTypeU{\ctx}{\EMV_3}{\TMV_2}
  }{
    \ctxSynTypeU{\ctx}{\EIf{\EMV_1}{\EMV_2}{\EMV_3}}{\TJoin{\TMV_1}{\TMV_2}}
  }{USIf}

  \judgment{
    \ctxAnaTypeU{\ctx}{\EMV_1}{\TBool} \\\\
    \ctxAnaTypeU{\ctx}{\EMV_1}{\TMV} \\
    \ctxAnaTypeU{\ctx}{\EMV_2}{\TMV}
  }{
    \ctxAnaTypeU{\ctx}{\ECIf{\EMV_1}{\EMV_2}{\EMV_3}}{\TMV}
  }{UAIf}
\end{mathpar}
%
In synthetic position, conditionals synthesize the \emph{join} of the branch types $\TMV_1$ and
$\TMV_2$, which we define inductively in \cref{fig:calculus-type-join}. We choose the ``more
specific'' type of the two. In analytic position, since
there is an expected type for both branches, we shift the blame for any errors to them.

\newcommand{\joinsTo}[3]{\ensuremath{\TJoin{#1}{#2} & = & #3}}
\begin{figure}[htbp]
  \raggedright
  \judgbox{\ensuremath{\TJoin{\TMV_1}{\TMV_2}}} is a \emph{partial} metafunction defined as follows:
  %
  \[\begin{array}{rcl}
    \joinsTo{\TUnknown}{\TMV}{\TMV} \\
    \joinsTo{\TMV}{\TUnknown}{\TMV} \\
    \joinsTo{\TNum}{\TNum}{\TNum} \\
    \joinsTo{\TBool}{\TBool}{\TBool} \\
    \joinsTo{(\TArrow{\TMV_1}{\TMV_2})}{(\TArrow{\TMV_1'}{\TMV_2'})}{\TArrow{(\TJoin{\TMV_1}{\TMV_1'})}{(\TJoin{\TMV_2}{\TMV_2'})}}
  \end{array}\]
  \vspace{-10px}
  \caption{Type join.}
  \label{fig:calculus-type-join}
\end{figure}

Following \textsc{USIf} and \textsc{UAIf}, we may derive the following marking and typing rules.
% Similar to previous cases, we recursively mark the guard and branches, constructing the marked
% conditional from the results.
%
\begin{mathpar}
  \judgment{
    \ctxAnaFixedInto{\ctx}{\EMV_1}{\ECMV_1}{\TBool} \\\\
    \ctxSynFixedInto{\ctx}{\EMV_2}{\ECMV_2}{\TMV_1} \\
    \ctxSynFixedInto{\ctx}{\EMV_3}{\ECMV_3}{\TMV_2}
  }{
    \ctxSynFixedInto{\ctx}{\EIf{\EMV_1}{\EMV_2}{\EMV_3}}{\ECIf{\ECMV_1}{\ECMV_2}{\ECMV_3}}{\TJoin{\TMV_1}{\TMV_2}}
  }{MKSIf}

  \judgment{
    \ctxAnaTypeM{\ctx}{\ECMV_1}{\TBool} \\\\
    \ctxSynTypeM{\ctx}{\ECMV_2}{\TMV_1} \\
    \ctxSynTypeM{\ctx}{\ECMV_3}{\TMV_2}
  }{
    \ctxSynTypeM{\ctx}{\ECIf{\ECMV_1}{\ECMV_2}{\ECMV_3}}{\TJoin{\TMV_1}{\TMV_2}}
  }{MSIf} \\

@@

However, in synthetic position, it may be the case that the two branch types have no join. This
occurs, in fact, when they are inconsistent, motivating the \emph{the inconsistent branches mark}
$\ECInconBr{\ECMV}{\ECMV}{\ECMV}$. Then, adding the following rules ensures totality on
conditionals:
%
\begin{mathpar}
  \judgment{
    \ctxAnaFixedInto{\ctx}{\EMV_1}{\ECMV_1}{\TBool} \\
    \ctxSynFixedInto{\ctx}{\EMV_2}{\ECMV_2}{\TMV_1} \\\\
    \ctxSynFixedInto{\ctx}{\EMV_3}{\ECMV_3}{\TMV_2} \\
    \inconsistent{\TMV_1}{\TMV_2}
  }{
    \ctxSynFixedInto{\ctx}{\EIf{\EMV_1}{\EMV_2}{\EMV_3}}{\ECInconBr{\ECMV_1}{\ECMV_2}{\ECMV_3}}{\TUnknown}
  }{MKSInconsistentBranches}

>>>>

Conditionals, however, present a more interesting case. In the unmarked language, we have both explicit
synthetic and analytic rules:
%
\begin{mathpar}
  \judgment{
    \ctxAnaTypeU{\ctx}{\EMV_1}{\TBool} \\\\
    \ctxSynTypeU{\ctx}{\EMV_2}{\TMV_1} \\
    \ctxSynTypeU{\ctx}{\EMV_3}{\TMV_2}
  }{
    \ctxSynTypeU{\ctx}{\EIf{\EMV_1}{\EMV_2}{\EMV_3}}{\TMeet{\TMV_1}{\TMV_2}}
  }{USIf}

  \judgment{
    \ctxAnaTypeU{\ctx}{\EMV_1}{\TBool} \\\\
    \ctxAnaTypeU{\ctx}{\EMV_1}{\TMV} \\
    \ctxAnaTypeU{\ctx}{\EMV_2}{\TMV}
  }{
    \ctxAnaTypeU{\ctx}{\ECIf{\EMV_1}{\EMV_2}{\EMV_3}}{\TMV}
  }{UAIf}
\end{mathpar}
%
In synthetic position, conditionals synthesize the \emph{meet} of the branch types $\TMV_1$ and
$\TMV_2$, which we define inductively in \cref{fig:calculus-type-meet}. We choose the ``more
specific'' type of the two. In analytic position, since
there is an expected type for both branches, we shift the blame for any errors to them.

\newcommand{\meetsTo}[3]{\ensuremath{\TMeet{#1}{#2} & = & #3}}
\begin{figure}[htbp]
  \raggedright
  \judgbox{\ensuremath{\TMeet{\TMV_1}{\TMV_2}}} is a \emph{partial} metafunction defined as follows:
  %
  \[\begin{array}{rcl}
    \meetsTo{\TUnknown}{\TMV}{\TMV} \\
    \meetsTo{\TMV}{\TUnknown}{\TMV} \\
    \meetsTo{\TNum}{\TNum}{\TNum} \\
    \meetsTo{\TBool}{\TBool}{\TBool} \\
    \meetsTo{(\TArrow{\TMV_1}{\TMV_2})}{(\TArrow{\TMV_1'}{\TMV_2'})}{\TArrow{(\TMeet{\TMV_1}{\TMV_1'})}{(\TMeet{\TMV_2}{\TMV_2'})}}
  \end{array}\]
  \vspace{-10px}
  \caption{Type meet.}
  \label{fig:calculus-type-meet}
\end{figure}

Following \textsc{USIf} and \textsc{UAIf}, we may derive the following marking and typing rules.
% Similar to previous cases, we recursively mark the guard and branches, constructing the marked
% conditional from the results.
%
\begin{mathpar}
  \judgment{
    \ctxAnaFixedInto{\ctx}{\EMV_1}{\ECMV_1}{\TBool} \\\\
    \ctxSynFixedInto{\ctx}{\EMV_2}{\ECMV_2}{\TMV_1} \\
    \ctxSynFixedInto{\ctx}{\EMV_3}{\ECMV_3}{\TMV_2}
  }{
    \ctxSynFixedInto{\ctx}{\EIf{\EMV_1}{\EMV_2}{\EMV_3}}{\ECIf{\ECMV_1}{\ECMV_2}{\ECMV_3}}{\TMeet{\TMV_1}{\TMV_2}}
  }{MKSIf}

  \judgment{
    \ctxAnaTypeM{\ctx}{\ECMV_1}{\TBool} \\\\
    \ctxSynTypeM{\ctx}{\ECMV_2}{\TMV_1} \\
    \ctxSynTypeM{\ctx}{\ECMV_3}{\TMV_2}
  }{
    \ctxSynTypeM{\ctx}{\ECIf{\ECMV_1}{\ECMV_2}{\ECMV_3}}{\TMeet{\TMV_1}{\TMV_2}}
  }{MSIf} \\

@@

However, in synthetic position, it may be the case that the two branch types have no meet. This
occurs, in fact, when they are inconsistent, motivating the \emph{the inconsistent branches mark}
$\ECInconBr{\ECMV}{\ECMV}{\ECMV}$. Then, adding the following rules ensures totality on
conditionals:
%
\begin{mathpar}
  \judgment{
    \ctxAnaFixedInto{\ctx}{\EMV_1}{\ECMV_1}{\TBool} \\
    \ctxSynFixedInto{\ctx}{\EMV_2}{\ECMV_2}{\TMV_1} \\\\
    \ctxSynFixedInto{\ctx}{\EMV_3}{\ECMV_3}{\TMV_2} \\
    \inconsistent{\TMV_1}{\TMV_2}
  }{
    \ctxSynFixedInto{\ctx}{\EIf{\EMV_1}{\EMV_2}{\EMV_3}}{\ECInconBr{\ECMV_1}{\ECMV_2}{\ECMV_3}}{\TUnknown}
  }{MKSInconsistentBranches}
